Child Signals of Need in Utila, Honduras

### 2024-08-23 Paper.qmd
Pared down analysis script and started adding tables and figures to the paper.

---
title: "Child Signals of Need in Utila, Honduras"
format: html
editor: visual
bibliography: signaling.bib
---

@@

We used 6-point Likert scales to get caregiver reports of 1) child signaling, 2) interpersonal conflict with each participating child, 3) and child alloparenting behavior. See Table 1 for exact wording.

@@

**Table 1**. Table showing variable names and question wording for measures about the frequency of behaviors. All questions had the following labels attached to each point on its Likert scale: Never; Once a month or less; More than once a month but less than once a week; More than once a week but not daily; Daily; Multiple times per day.

As these responses reflect frequencies of actual behavior, we converted these to counts which reflected the number of times per month a child engaged in these behaviors. The original wording of the Likert scale and the frequency conversion are shown in Table 2.

@@

**Table 2.** Table showing conversions from Likert scale to count value for the frequencies of child behaviors included in this study. Variables created with this scheme include SadFreqN, CryFreqN, TantrumFreqN, RunawayFreqN, ConflictFreqN, and AlloparentingFreqN.

@@

We also used 6-point Likert scales to get parental reports of 1) the relative neediness of each participating child and 2) the relative investment they provide to each participating child compared to the rest of their children. See Table 3 for exact wording.

@@

**Table 3.** Table showing variable names and question wording for measures about relative need and investment. Both questions had the following labels attached to each point on its Likert scale: Much less; Less; The same amount; More; Much more.

@@

Caretakers were also asked three questions using a five point Likert scale about their child's current health which involved the likelihood of them becoming sick. We then inverted the scores for two measures so that a score of 5 corresponded with the lowest susceptibility to illness before averaging the scores of these to create the CurrentHealthMean variable. Means were still calculated if 1-2 values had an NA (only xxx cases), with the numerator being the number of variables with data. See Table 4 for the wording of these variables.

@@

: **Table 4.** Table showing variable names and question wording for measures about a child's current likelihood of becoming sick. All three questions were asked using a Likert scale from 1-5 where 1 was labeled "Strongly disagree" and 5 was labeled "Strongly agree." The scores for Currenthealth1 and Currenthealth2 were inverted so that 5 corresponded with the lowest susceptibility to illness.

@@

```{r sumtab_children}
#| echo: false
#| label: sumtab_children
#| fig-cap: "Summary statistics for child study variables."

vars_c <- c(
  ChildAge = "Age (years)",
  SadFreqN = "Sad frequency (times per month)",
  CryFreqN = "Crying frequency (times per month)",
  TantrumFreqN = "Temper tantrum frequency (times per month)",
  AlloparentingFreqN = "Frequency of child's own alloparetning effort (times per month)",
  ConflictFreqN = "Frequency of conflict with focal caretaker (times per month)",
  SignalFreq = "Frequency of signals summed (times per month)",
  SignalCost = "Signal cost",
  SignalFreqMax = "Frequency of most common signal (times per month)",
  Currenthealth1 = "Currenthealth1",
  Currenthealth2 = "Currenthealth2",
  Currenthealth3 = "Currenthealth3",
  CurrentHealthMean = "Mean of current health measures"
)

@@

```{r sumtab_adults}
#| echo: false
#| label: sumtab_cadults
#| fig-cap: "Summary statistics for adult and household study variables."

@@

>>>>

---
title: "Child Signals of Need in Utila, Honduras"
execute: 
  echo: false
  warning: false
format: html
editor: visual
tbl-cap-location: bottom
bibliography: signaling.bib
---

@@

We used 6-point Likert scales to get caregiver reports of 1) child signaling, 2) interpersonal conflict with each participating child, 3) and child alloparenting behavior. See @tbl-freqvars for exact wording.

@@

: Table showing variable names and question wording for measures about the frequency of behaviors. All questions had the following labels attached to each point on its Likert scale: Never; Once a month or less; More than once a month but less than once a week; More than once a week but not daily; Daily; Multiple times per day. {#tbl-freqvars}

As these responses reflect frequencies of actual behavior, we converted these to counts which reflected the number of times per month a child engaged in these behaviors. The original wording of the Likert scale and the frequency conversion are shown in @tbl-convertedfreqs.

@@

: Table showing conversions from Likert scale to count value for the frequencies of child behaviors included in this study. Variables created with this scheme include SadFreqN, CryFreqN, TantrumFreqN, RunawayFreqN, ConflictFreqN, and AlloparentingFreqN. {#tbl-convertedfreqs}

@@

We also used 6-point Likert scales to get parental reports of 1) the relative neediness of each participating child and 2) the relative investment they provide to each participating child compared to the rest of their children. See Table @tbl-relvars for exact wording.

@@

: Table showing variable names and question wording for measures about relative need and investment. Both questions had the following labels attached to each point on its Likert scale: Much less; Less; The same amount; More; Much more. {#tbl-relvars}

@@

Caretakers were also asked three questions using a five point Likert scale about their child's current health which involved the likelihood of them becoming sick. We then inverted the scores for two measures so that a score of 5 corresponded with the lowest susceptibility to illness before averaging the scores of these to create the CurrentHealthMean variable. Means were still calculated if 1-2 values had an NA (only xxx cases), with the numerator being the number of variables with data. See Table @tbl-healthvars for the wording of these variables.

@@

: Table showing variable names and question wording for measures about a child's current likelihood of becoming sick. All three questions were asked using a Likert scale from 1-5 where 1 was labeled "Strongly disagree" and 5 was labeled "Strongly agree." The scores for Currenthealth1 and Currenthealth2 were inverted so that 5 corresponded with the lowest susceptibility to illness. {#tbl-healthvars}

@@

```{r sumtab_children}
#| label: tbl-sumtab_children
#| tbl-cap: Summary statistics for child study variables.
vars_c <- c(
  ChildAge = "Age (years)",
  SadFreqN = "Sad frequency (times per month)",
  CryFreqN = "Crying frequency (times per month)",
  TantrumFreqN = "Temper tantrum frequency (times per month)",
  AlloparentingFreqN = "Frequency of child's own alloparetning effort (times per month)",
  ConflictFreqN = "Frequency of conflict with focal caretaker (times per month)",
  SignalFreq = "Frequency of signals summed (times per month)",
  SignalCost = "Signal cost",
  SignalFreqMax = "Frequency of most common signal (times per month)",
  Currenthealth1 = "Currenthealth1",
  Currenthealth2 = "Currenthealth2",
  Currenthealth3 = "Currenthealth3",
  CurrentHealthMean = "Mean of current health measures"
)

@@

```{r sumtab_adults}
#| label: tbl-sumtab_adults
#| tbl-cap: Summary statistics for adult and household study variables.

@@

## Distributions of signaling behavior, alloparenting effort, and conflict

:::{#fig-barplot_SignalFreq}
```{r}
barplot_SignalFreq
```

caption
:::

:::{#fig-barplot_ConflictFreq}
```{r}
barplot_conflict
```

caption
:::

:::{#fig-barplot_AlloparentingFreq}
```{r}
barplot_alloparenting
```

caption
:::

## Signaling behavior is negatively associated with age

```{r}
#| label: tbl-main_models
#| tbl-cap: Regression tables for our main exploratory models.
r_vars <- c(
  ChildAge = "Age (years)",
  SexMale = "Sex (male)",
  OtherChildrenHH = "Other children",
  LogIncome = "Income (log)",
  number_adults = "Adults in home",
  PartnerStatusUnpartnered = "Partner status (single)",
  EducationLevelYears = "Parent ed. (years)",
  ConflictFreqN = "Conflict freq.",
  SignalFreq = "Signal freq.",
  SignalCost = "Signal cost",
  SignalFreqMax = "Max signal freq.",
  CurrentHealthMean = "Percieved immunity",
  SadFreqN = "Sadness freq.",
  CryFreqN = "Crying freq.",
  TantrumFreqN = "Tantrum freq.",
  AlloparentingFreqN = "Child allocare"
)
modelsummary(list(Sadness = msadH, Crying = mcryH, Tantrum = mtantrumH, SignalFreq = msfH, SignalCost = mscH, SignalFreqMax = mmsfH), gof_map = "nobs", coef_rename = r_vars, stars = TRUE, coef_omit = "(Intercept)")
```

:::{#fig-effect1}
```{r}
plot_predictions(msadH, condition = "ChildAge", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Frequency of sadness (per month)") + xlab("Child age (years)") + xlim(5,20)
```

The effect of the age on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior. Bars are 95% CIs (only taking the uncertainty of fixed-effect parameters into account). Each dot is 1 child. Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate). See @tbl-main_models for the corresponding regression table.
:::

:::{#fig-effect2}
```{r}
plot_predictions(mcryH, condition = "ChildAge", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Crying frequency (per month)") + xlab("Child age (years)") + xlim(5,20)
```

combine with other age plots and use caption 1
:::

:::{#fig-effect3}
```{r}
plot_predictions(mtantrumH, condition = "ChildAge", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Tantrum frequency (per month)") + xlab("Child age (years)") + xlim(5,20)
```

combine with other age plots and use caption 1
:::

:::{#fig-effect4}
```{r}
plot_predictions(mscH, condition = "ChildAge", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Summed signal cost") + xlab("Child age (years)") + xlim(5,20)
```

combine with other age plots and use caption 1
:::

:::{#fig-effect5}
```{r}
plot_predictions(msfH, condition = "ChildAge", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Summed frequency of signals") + xlab("Child age (years)") + xlim(5,20)
```

combine with other age plots and use caption 1
:::

:::{#fig-effect6}
```{r}
plot_predictions(mmsfH, condition = "ChildAge", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Frequency of most common signal") + xlab("Child age (years)") + xlim(5,20)
```

combine with other age plots and use caption 1
:::

## Signaling behavior is positively associated with conflict with primary caretakers

:::{#fig-effect1c}
```{r}
plot_predictions(msadH, condition = "ConflictFreqN", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Frequency of sadness (per month)") + xlab("Frequency of conflict (per month)") 
```

The effect of the frequency of conflict with one's primary caretaker on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior. Bars are 95% CIs (only taking the uncertainty of fixed-effect parameters into account). Each dot is 1 child. Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate). See @tbl-main_models for the corresponding regression table.
:::

:::{#fig-effect2c}
```{r}
plot_predictions(mcryH, condition = "ConflictFreqN", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Crying frequency (per month)") + xlab("Frequency of conflict (per month)")
```

combine with other confict plots and use caption 1
:::

:::{#fig-effect3c}
```{r}
plot_predictions(mtantrumH, condition = "ConflictFreqN", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Tantrum frequency (per month)") + xlab("Frequency of conflict (per month)")
```

combine with other conflict plots and use caption 1
:::

:::{#fig-effect4c}
```{r}
plot_predictions(mscH, condition = "ConflictFreqN", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Summed signal cost") + xlab("Frequency of conflict (per month)")
```

combine with other conflict plots and use caption 1
:::

:::{#fig-effect5c}
```{r}
plot_predictions(msfH, condition = "ConflictFreqN", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Summed frequency of signals") + xlab("Frequency of conflict (per month)") 
```

combine with other conflict plots and use caption 1
:::

:::{#fig-effect6c}
```{r}
plot_predictions(mmsfH, condition = "ConflictFreqN", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Frequency of most common signal") + xlab("Frequency of conflict (per month)")
```

combine with other conflict plots and use caption 1
:::

## Interaction between child sex and child alloparenting

## Summed signal cost and tantrum frequency are negatively associated with body fat

```{r}
#| label: tbl-bodyfat_models
#| tbl-cap: Regression tables for the models examining body fat measures.
r_vars <- c(
  ChildAge = "Age (years)",
  SexMale = "Sex (male)",
  OtherChildrenHH = "Other children",
  LogIncome = "Income (log)",
  number_adults = "Adults in home",
  PartnerStatusUnpartnered = "Partner status (single)",
  EducationLevelYears = "Parent ed. (years)",
  ConflictFreqN = "Conflict freq.",
  SignalFreq = "Signal freq.",
  SignalCost = "Signal cost",
  SignalFreqMax = "Max signal freq.",
  CurrentHealthMean = "Percieved immunity",
  SadFreqN = "Sadness freq.",
  CryFreqN = "Crying freq.",
  TantrumFreqN = "Tantrum freq.",
  AlloparentingFreqN = "Child allocare",
  TricepR = "Tricep skinfold (residuals)",
  SubscapR = "Subscapular skinfold (residuals)",
  BodyFat = "BodyFat (composite)"
)
modelsummary(list(SignalCost = mBFc, Tantrum = mBFtantrum, SignalCost = msubC, Tantrum = msubTantrum, SignalCost = mtriC, Tantrum = mtriTantrum), gof_map = "nobs", coef_rename = r_vars, stars = TRUE, coef_omit = "(Intercept)")
```

:::{#fig-cost_bodyfat}
```{r}
plot_predictions(mBFc, condition = "BodyFat", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Summed signal cost") + xlab("Body fat composite") + xlim(0,10) + ylim(0,200)
```

combine with other age plots and use caption 1
:::

:::{#fig-tantrum_bodyfat}
```{r}
plot_predictions(mBFtantrum, condition = "BodyFat", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Frequency of tantrums (per month)") + xlab("Body fat composite") + xlim(0,10)
```

combine with other body fat plots and use caption 1
:::

## Cargivers percieve older children as relatively needy

## Parental responses to last signal of need

## Sociodemographics
